Demography and linked selection in wild and domesticated maize

### 2015-08-12 manuscript.tex

To ensure that this analysis is resilient to the intricacies of our sample, we investigated diversity surrounding the maize gene \emph{tga1}, one of the few known instances of a hard sweep corresponding to a causal missense substitution in maize \cite{wang2015}. In the case of this known sweep, we observed a reduction in diversity surrounding the causative substitution to levels much lower than seen for synonymous substitutions (Figure S1) \jri{please add refs to link to supp. figs if possible}, confirming that this method should be informative if hard sweeps were the prevailing mode of selection in maize. Still, to ensure that the approach is robust to potential shortcomings such as limited power in the face of background selection \cite{enard2014}, we sought to determine if patterns of diversity surrounding synonymous and non-synonymous substitutions differ when we restrict our analysis to the most- or least-conserved genes in maize. When we re-analyzed subsets of genes with the 10\% highest and 10\% lowest genomic evolutionary rate profile (GERP) score, putatively corresponding to genes undergoing the strongest and weakest purifying selection \cite{davydov2010, cooper2005, rodgers2015}, we again saw no significant difference in diversity surrounding synonymous and non-synonymous sites (Figure S2).

\begin{figure*}[!htb]
%\vspace*{.05in}
\centering
\includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_Significance_June}
\hspace{0.05\textwidth} \includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvT_Folded2_Significance_June}
\caption{Pairwise diversity surrounding synonymous and non-synonymous
  substitutions in maize and teosinte. \jri{a bit too much vertical white space} {\bf A:} Maize diversity; {\bf B:} Teosinte
diversity. Bootstrap-based 95\% confidence intervals are depicted via shading. Inset plots depict a larger range on the x-axis. \label{fig:hardSweeps}}
\end{figure*}

%%%%
%%%% This outline no longer needed in-text because results headers follow
%%%%
% \jri{
% {\bf Outline}
% \begin{enumerate}
% \item Diversity differs between genes, not genes
% \item No hard sweeps
% \item soft sweeps? (putting it here means if we find a lot we may have to redo stuff below on the subset of genes w/o evidence of soft sweeps.)
% \item purifying selection differs ($\pi$)
% \item Estimating the bneck
% \item purifying selection (singletons)
% \item fit the B curve
% \item simulations
% \end{enumerate}
% }

@@

\subsection{Plant materials}
We made use of published sequence of inbred accessions of teosinte (Z. \emph{mays} ssp. \emph{parviglumis}) and maize landraces from the Maize HapMap2 panel \cite{chia2012, lemmon2014}. 
We removed \X teosinte and \X maize individuals that appeared as outliers in an initial principal component analysis  (Figure S5), leaving 13 teosinte and 23 maize that were used for all subsequent analyses (Table S1).

>>>>

To ensure that this analysis is resilient to the intricacies of our sample, we investigated diversity surrounding the maize gene \emph{tga1}, one of the few known instances of a hard sweep corresponding to a causal missense substitution in maize \cite{wang2015}. In the case of this known sweep, we observed a reduction in diversity surrounding the causative substitution to levels much lower than seen for synonymous substitutions (Figure \ref{sFig:tga1}), confirming that this method should be informative if hard sweeps were the prevailing mode of selection in maize. Still, to ensure that the approach is robust to potential shortcomings such as limited power in the face of background selection \cite{enard2014}, we sought to determine if patterns of diversity surrounding synonymous and non-synonymous substitutions differ when we restrict our analysis to the most- or least-conserved genes in maize. When we re-analyzed subsets of genes with the 10\% highest and 10\% lowest genomic evolutionary rate profile (GERP) score, putatively corresponding to genes undergoing the strongest and weakest purifying selection \cite{davydov2010, cooper2005, rodgers2015}, we again saw no significant difference in diversity surrounding synonymous and non-synonymous sites (Figure \ref{sFig:consUncons}).

\begin{figure*}[!htb]
%\vspace*{.05in}
\centering
\includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_Significance_Aug}
\hspace{0.05\textwidth} \includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvT_Folded2_Significance_Aug}
\caption{Pairwise diversity surrounding synonymous and non-synonymous
  substitutions in maize and teosinte. {\bf A:} Maize diversity; {\bf B:} Teosinte
diversity. Bootstrap-based 95\% confidence intervals are depicted via shading. Inset plots depict a larger range on the x-axis. \label{fig:hardSweeps}}
\end{figure*}

@@

\subsection{Plant materials}
We made use of published sequence of inbred accessions of teosinte (Z. \emph{mays} ssp. \emph{parviglumis}) and maize landraces from the Maize HapMap2 panel \cite{chia2012, lemmon2014}. 
We removed \X teosinte and \X maize individuals that appeared as outliers in an initial principal component analysis  (Figure \ref{sFig:PCA}), leaving 13 teosinte and 23 maize that were used for all subsequent analyses (Table sTab:list}).
